---
title: "A Quick Look into Bike Buyers Dataset"
author: "Florencia"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
#import csv
bike_buyers <- read.csv("E:/Bike Buyers/bike_buyers.csv")
```

### 1.Basic data characteristics
```{r}
dim(bike_buyers)
```
**EXPLANATION**

The dim function returns the dimension of the bike_buyers dataset. it shows that the bike_buyers dataset has 1000 rows and 13 columns ( 1000 instances and 13 attributes)


```{r}
str(bike_buyers)
writeLines("\n")
sapply(bike_buyers, class)
```
**EXPLANATION**

there are two different data types in the bike_buyers dataset which are integer (ID, Income, Children, Cars, Age) and character (Marital.Status, Gender, Education, Occupation, Home.Owner, Commute Distance, Region, Purchased.Bike)

```{r}
BasicSummary <- function(df, dgts = 3){
  m <- ncol(df)
varNames <- colnames(df)
varType <- vector("character",m)
topLevel <- vector("character",m)
topCount <- vector("numeric",m)
missCount <- vector("numeric",m)
levels <- vector("numeric", m)

for (i in 1:m){
x <- df[,i]
varType[i] <- class(x)
xtab <- table(x, useNA = "ifany")
levels[i] <- length(xtab)
nums <- as.numeric(xtab)
maxnum <- max(nums)
topCount[i] <- maxnum
maxIndex <- which.max(nums)
lvls <- names(xtab)
topLevel[i] <- lvls[maxIndex]
missIndex <- which((is.na(x)) | (x == "") | (x == " "))
missCount[i] <- length(missIndex)
}
n <- nrow(df)
topFrac <- round(topCount/n, digits = dgts)
missFrac <- round(missCount/n, digits = dgts)
## #
summaryFrame <- data.frame(variable = varNames, type = varType,
 levels = levels, topLevel = topLevel,
 topCount = topCount, topFrac = topFrac,
 missFreq = missCount, missFrac = missFrac)
 return(summaryFrame)
 }

BasicSummary(bike_buyers)
```
 **EXPLANATION**
 
 It is clear that all the variables have clear and simple explanatory names which are not difficult to understand and it describes the data in the dataset.
 From the 7 variables in the dataset, 5 of them were integer, and the rest is character.
 It can be seen that the integer variables has more levels than the character variables.
 
 
### 2. Summary Statistics

```{r}
summary(bike_buyers)

writeLines("Mean:")
sapply(bike_buyers[, c(4,12)], mean, na.rm=TRUE)
writeLines("\nDescription:")
sapply(bike_buyers[, c(4,12)], quantile, na.rm=TRUE)
```

```{r}
library(Hmisc)
describe(bike_buyers)
```
 **EXPLANATION**
 
 Missing values were found in the bike_buyers dataset, 7 missing values in Marital.Status variable, 11 in Gender variable, 6 in Income variable, 8 missing values iin CHildren variable, 4 in Home.Owner variable, 9 in Cars variables, and 8 missing values in the Age variable.
 
```{r}
bike_buyers[, c(2,3,6:8, 10, 11, 13)] <- lapply(bike_buyers[, c(2,3,6:8, 10, 11, 13)], as.factor)
```


### 3. Data anomalies
```{r}
ThreeSigma <- function(x, t = 3){

 mu <- mean(x, na.rm = TRUE)
 sig <- sd(x, na.rm = TRUE)
 if (sig == 0){
 message("All non-missing x-values are identical")
}
 up <- mu + t * sig
 down <- mu - t * sig
 out <- list(up = up, down = down)
 return(out)
 }

Hampel <- function(x, t = 3){

 mu <- median(x, na.rm = TRUE)
 sig <- mad(x, na.rm = TRUE)
 if (sig == 0){
 message("Hampel identifer implosion: MAD scale estimate is zero")
 }
 up <- mu + t * sig
 down <- mu - t * sig
 out <- list(up = up, down = down)
 return(out)
 }
   
BoxplotRule<- function(x, t = 1.5){

 xL <- quantile(x, na.rm = TRUE, probs = 0.25, names = FALSE)
 xU <- quantile(x, na.rm = TRUE, probs = 0.75, names = FALSE)
 Q <- xU - xL
 if (Q == 0){message("Boxplot rule implosion: interquartile distance is zero")
 }
 up <- xU + t * Q
 down <- xU - t * Q
 out <- list(up = up, down = down)
 return(out)
}   

ExtractDetails <- function(x, down, up){

 outClass <- rep("N", length(x))
 indexLo <- which(x < down)
 indexHi <- which(x > up)
 outClass[indexLo] <- "L"
 outClass[indexHi] <- "U"
 index <- union(indexLo, indexHi)
 values <- x[index]
 outClass <- outClass[index]
 nOut <- length(index)
 maxNom <- max(x[which(x <= up)])
 minNom <- min(x[which(x >= down)])
 outList <- list(nOut = nOut, lowLim = down,
 upLim = up, minNom = minNom,
 maxNom = maxNom, index = index,
 values = values,
 outClass = outClass)
 return(outList)
 }
```

```{r}
FindOutliers <- function(x, t3 = 3, tH = 3, tb = 1.5){
 threeLims <- ThreeSigma(x, t = t3)
 HampLims <- Hampel(x, t = tH)
 boxLims <- BoxplotRule(x, t = tb)

 n <- length(x)
 nMiss <- length(which(is.na(x)))

 threeList <- ExtractDetails(x, threeLims$down, threeLims$up)
 HampList <- ExtractDetails(x, HampLims$down, HampLims$up)
 boxList <- ExtractDetails(x, boxLims$down, boxLims$up)

 sumFrame <- data.frame(method = "ThreeSigma", n = n,
 nMiss = nMiss, nOut = threeList$nOut,
 lowLim = threeList$lowLim,
 upLim = threeList$upLim,
 minNom = threeList$minNom,
 maxNom = threeList$maxNom)
 upFrame <- data.frame(method = "Hampel", n = n,
 nMiss = nMiss, nOut = HampList$nOut,
 lowLim = HampList$lowLim,
 upLim = HampList$upLim,
 minNom = HampList$minNom,
 maxNom = HampList$maxNom)
 sumFrame <- rbind.data.frame(sumFrame, upFrame)
 upFrame <- data.frame(method = "BoxplotRule", n = n,
 nMiss = nMiss, nOut = boxList$nOut,
 lowLim = boxList$lowLim,
 upLim = boxList$upLim,
 minNom = boxList$minNom,
 maxNom = boxList$maxNom)
 sumFrame <- rbind.data.frame(sumFrame, upFrame)

 threeFrame <- data.frame(index = threeList$index,
 values = threeList$values,
 type = threeList$outClass)
 HampFrame <- data.frame(index = HampList$index,
 values = HampList$values,
 type = HampList$outClass)
 boxFrame <- data.frame(index = boxList$index,
 values = boxList$values,
 type = boxList$outClass)
 outList <- list(summary = sumFrame, threeSigma = threeFrame,
 Hampel = HampFrame, boxplotRule = boxFrame)
 return(outList)
}


```

```{r}
FullSummary <- FindOutliers(bike_buyers$Income)
FullSummary$summary
```
 **EXPLANATION**
 From these three method of finding the outliers, three of them detect the same amount of the outliers which is 10 outliers.
 For the upper and lower limit, the BoxplotRule has the lowest upper and lower outlier limit among the three of them, but it does'nt give that big/ much difference
 The lower and upper limits of the non-outlying data values of the three rule has the same value

```{r}
rcorr(as.matrix(bike_buyers[c(1,4,5, 9, 12)]), type = "spearman")
```
```{r}
hist.data.frame(bike_buyers)
plot(bike_buyers)
```
```{r}
Table <- table(bike_buyers$Income, bike_buyers$Purchased.Bike, bike_buyers$Gender)
print(Table)
```


```{r}
Table <- table(bike_buyers$Age, bike_buyers$Marital.Status)
print(Table)
```
```{r}
matrix <- layout( matrix(c(1,2,3,4), nrow=2, byrow=TRUE) )

mosaicplot(Gender~Purchased.Bike, 
           data = bike_buyers,
           main = "Gender vs Purchased Bike",
           col = "pink",
           las=1,
           shade = TRUE)

boxplot(Income~Purchased.Bike,
        data = bike_buyers,
        xlab = "Purchased Bike",
        main = "Purchased bike status over Income",
        col = "lightblue")

boxplot(Age~Marital.Status,
           data = bike_buyers,
        main = "Marital Status by age",
        col = "lightgreen")

mosaicplot(Children~Purchased.Bike,
           data = bike_buyers,
           main = "",
           col = "lightyellow")
```
 **EXPLANATION**
 
The first plot in the upper left tells us that the amount of bike purchased by the female and male gender has not much difference
 
In the second plot (the upper right), the income of the buyers doesn't really affect the purchased bike, so people with higher income will not be guaranteed to buy the bike.

The third plot (lower left), indicates that most people with high age  have the status of being married and for the last plot, people with range 0-4 children tend to purchased bike rather than people with 5 children which is the most amount of children in the dataset.
 
